Effect of multiple reflection of the sky signal off the HERA element on the measurements of 21cm power spectrum

### 2015-12-04 DishReflectometry/Paper1_v0.tex
Continued rewriting section 4

\subsection{Effect of first order reflections on visibility and delay spectrum}
Upon first incidence of the sky signal $\volt_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, the net voltage at the output of the two antennas could be written as,    
\begin{eqnarray}
\volt_{1} & = & a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp  (1-\Gamma_{f})^{2}\nonumber\\
         & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber\\
\label{eq4}
\end{eqnarray}
\begin{eqnarray}
\volt_{2} & = & \left [ a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp (1-\Gamma_{f})^{2}\right ] \fngexp \nonumber\\
         & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} \dfngexp  \right] \fngexp \nonumber\\
\label{eq5}
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna field pattern $a$ and denote them as $a_{f}= a(1-\Gamma_{f})$ here after and define the antenna cross power pattern as $\bm_{f}(\thhat,\nu)=\bmvolt_{f}(\thhat,\nu)\bmvolt_{f}^{*}(\thhat,\nu)$ . Hence, the complex visibility could be,

\begin{eqnarray}
\vis(\vec b, \nu) & = & \int \volt_{1}\volt_{2}^{*} d\Omega \nonumber\\
		        & = & \int \bm_{f} \volt_{sky}|^{2} |\left [ 1+(1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} \dfngexp  \right]|^{2} \fngexp  d\Omega
\label{eq6}		        
\end{eqnarray}

 Furthermore, combining all the terms that contain instrument response we define a common instrument parameter, written as, 
\begin{eqnarray}
\Sigma(\vec b, \nu) & = & \bm_{f} |\left [ 1+ (1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]|^{2}\nonumber\\
 	    & = &  \bm_{f}  \left[ 1+ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} + 2 \Re \left ( (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp \right) \right] 
\label{eq{7}}	    
\end{eqnarray}

@@

\subsection{Effect of multiple reflections on the delay spectrum} 
Correlation between the higher order reflections of the signal in any antenna with the direct component of the other would result in a delay response at integral multiple of $\Delta \tau$ and vice versa. 
If $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{equation}
\volt_{1} =  a_{f} \volt_{sky}[1+ (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp + (1-\Gamma_{f})^{2} (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ ....+ (1-\Gamma_{f})^{n}(\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}]
\end{equation}
and,

\begin{eqnarray}
\volt_{2} & = &  a_{f} \volt_{sky}[1+ (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp + (1-\Gamma_{f})^{2} (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&& ....+ (1-\Gamma_{f})^{n}(\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \ast \fngexp 
\end{eqnarray}

@@

\begin{eqnarray}
\Sigma(\vec b, \nu) & = & a_{f}  [ 1+ 2 \Re [(1-\Gamma_{f}) \Gamma_{f}\Gamma_{d} \dfngexp+ (1-\Gamma_{f})^{2}(\Gamma_{f}\Gamma_{d})^2(\dfngexp)^{2}+ \nonumber\\
&&  ....+(1-\Gamma_{f})^{n} (\Gamma_{f}\Gamma_{d})^{n}(\dfngexp)^{n} ] ]
\end{eqnarray}

@@

\begin{eqnarray}\label{eqn:ratio1}
{I_{meas} \over I_{sky} } & = & \Sigma(\vec b, \nu) \nonumber\\
      & = & a_{f}  [ 1+ 2 \Re [(1-\Gamma_{f}) \Gamma_{f}\Gamma_{d} \dfngexp+ (1-\Gamma_{f})^{2}(\Gamma_{f}\Gamma_{d})^2(\dfngexp)^{2}+ \nonumber\\
&&  ....+(1-\Gamma_{f})^{n} (\Gamma_{f}\Gamma_{d})^{n}(\dfngexp)^{n} ] ]
\end{eqnarray}

The right hand side of this equation is the instrument response kernel of a single HERA element as well as a representative of the instrument response in the measured visibility data between a pair of antennas. Our goal is to measure this kernel as a function of frequency and determine its effect on the delay spectrum of the visibility. 
\subsection{Method}

In our experimental set-up, we use the feed as a transmitter and transmit a pulse. If the initial pulse is a broadband signal,
$I_{tr}$, sent to the feed via a $75ft$ cable by a vector network
analyser (VNA), a delay domain measurement of the system is accomplished by
taking the Fourier transform of the complex return loss of the feed measured by VNA. When the signal is incident on
the feed, part of the incident power ($\Gamma_{f}$) is reflected back to the
measuring device (dashed red arrows in Figure \ref{fig:cartoon}) and
$(1-\Gamma_{f})$ is radiated by the feed (solid red arrows in Figure
\ref{fig:cartoon}) which illuminates the dish. While most of the incident signal is reflected back into space by the dish, the reflected signal from the dish vertex returns to
the feed. This incident signal is now
reflected back and forth in between the feed and the dish much like the sky
signal reflection discussed previously.  Hence, if $I_{r}$ is the power
incident back on the feed for the first time then the reflected power, $I_{refl}$,
back into the VNA would be:

\begin{equation}\label{eqn:series2}
I_{refl} =  I_{r}A_{f}[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

Once again, note that we consider one reflection from the feed and its subsequent reflection from the dish as one reflection in total. Equation \ref{eqn:series2} is similar to Equation \ref{eqn:series1}, with different incident powers.

Recall that $I_{r}$ is the initial power that is incident back on the feed, which is just the feed radiated power reflected off the dish:

@@

Also note that the first reflection of the signal sent by the VNA occurs at the feed end. Hence the total returned power $P_{ret}$, to the VNA  would be:

@@

  \begin{eqnarray}\label{eqn:ratio2}
 {I_{ret} \over I_{tr} } & = & \Gamma_{f}
  +  \Gamma_{d}(1-\Gamma_f)^{2} [e^{i\phi}+ \Gamma_{f}\Gamma_{d} e^{i2\phi}+  ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{i n\phi}]\nonumber\\
  & = & \Gamma_{f} + { (1-\Gamma_f)^{2}\over \Gamma_{f} } \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i\phi}]^{n}
   \nonumber\\
\end{eqnarray}

@@

The VNA measures the magnitude and phase of the quantity ${I_{ret}\over I_{tr}}$
as a function of frequency as shown in Figure \ref{fig:freq}. In our measurement
set-up, the first reflection occurs at the feed $\Gamma_{f}$, so
$({I_{ret} \over I_{tr} }  - \Gamma_{f}) $ gives an estimate of the delay
spectrum of the sky signal. In delay domain, the relative signal strength at
zero delay represents the factor $\Gamma_{f}$ while the signal strength at any
other delay represents any delayed signal that enters the feed after being
reflected from the feed surroundings.

\section{Methodology}{\label{sec:methods}}

Our reflectometry measurements are made using a prototype HERA dish at NRAO in Green Bank, WV. The dish is a $14\,m$ diameter
parabolic reflector structurally supported with 3 telephone poles. The
reflective material is made up of wire mesh that is attached to PVC
pipes, forming the parabolic shape of the dish. With the current iteration of
the HERA dish, the feed consists of a PAPER dipole encased in a cylindrical cage
encompassing the backplane. The PAPER feed and the backplane (which is aimed at
preventing feed-to-feed interaction between neighboring dishes) is raised and
lowered by a three-pulley system. The focal height of the dish is $4.5m$
($\sim{14.76}$ft).

%The heights measured in our experiment were made from
%the balun to the top of the concrete hub. However, the focal height should be
%measured from the backplane to point where the wire mesh would intersect the
%concrete hub at the vertex of the dish. To account for this height discrepency,
%we add 1.96 ft to all of our measured heights. These are the heights quoted in
%the plots.[XXX not done yet but will be done]
Our measurements are made with a FieldFox in VNA mode. In this mode, a pulse
is generated in the FieldFox and sent through a $75ft$ $50\Omega$ cable that
connects to the feed with a 4:1 passive balun. The magnitude and phase of the complex return loss is measured between 50 to $500MHz$ over 1024 frequency channels which gives a frequency resolution of 0.44MHz. The delay resolution of the measurements is $\Delta{t}=2.22ns$.
In addition, we note that the round trip of a reflection from the feed to the dish
is $9m$, which corresponds to a delay of $30ns$.

@@

%need to add in third : Antenna feed focal length 4.5 m which corresponds to a
%signal propagation delay of 0.3 ns. Hence, subsequent reflections from the dish
%vertex are expected to be at a delay which is integral multiple of 0.3 nS.
\begin{figure}[ht!]
\centering
\includegraphics[totalheight=0.4\textheight]{plots/frequency_amp_phase_fullbw.png}
\caption{Amplitude and phase of the measured return loss. Colored dashed lines
mark three different frequency bands: $140-160MHz$ (typical power spectrum bandwidth), $100-200MHz$ (PAPER bandwidth), and
$50-250MHz$ (HERA bandwidth).}
\label{fig:freq}
\end{figure}

Frequency domain data, as shown in Figure \ref{fig:freq}, can be Fourier-transformed to compute the return loss in the delay domain. Because the measured data is band-limited between $50$ to $500MHz$, it is analogous to multiply the measured data by a square window function. Then in delay domain, the response would be convolved with a $sinc$ function. This may result in excess power at high
delays due to the sidelobes of the $sinc$ function. Hence, appropriate windowing of the measured data set is necessary before taking the Fourier transform.
We have chosen a Hamming window for our analysis, and its effectiveness of this
window function compared to others is illustrated in Figure
\ref{fig:window}.

>>>>

\subsection{Effect of first order reflections on visibility and delay spectrum}
Upon first incidence of the sky signal $\volt_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, the net voltage at the output of the two antennas could be written as,    
\begin{eqnarray}
\volt_{1} & = & a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp  (1-\Gamma_{f})^{2}\nonumber\\
         & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber
\end{eqnarray}
\begin{eqnarray}
\volt_{2} & = & \left [ a \volt_{sky}(1-\Gamma_{f}) + a \volt_{sky} \Gamma_{f} \Gamma_{d} \dfngexp (1-\Gamma_{f})^{2}\right ] \fngexp \nonumber\\
         & = & a \volt_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} \dfngexp  \right] \fngexp \nonumber
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna field pattern $a$ and denote them as $a_{f}= a(1-\Gamma_{f})$ here after and define the antenna cross power pattern as $\bm_{f}(\thhat,\nu)=\bmvolt_{f}(\thhat,\nu)\bmvolt_{f}^{*}(\thhat,\nu)$ . Hence, the complex visibility could be,

\begin{eqnarray}
\vis(\vec b, \nu) & = & \int \volt_{1}\volt_{2}^{*} d\Omega \nonumber\\
		        & = & \int \bm_{f} \volt_{sky}|^{2} |\left [ 1+(1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} \dfngexp  \right]|^{2} \fngexp  d\Omega
\label{eq4}		        
\end{eqnarray}

 Furthermore, combining all the terms that contain instrument response we define a common instrument parameter, written as, 
\begin{eqnarray}
\Sigma(\vec b, \nu) & = & \bm_{f} |\left [ 1+ (1-\Gamma_{f}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]|^{2}\nonumber\\
 	    & = &  \bm_{f}  \left[ 1+ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} + 2 \Re \left ( (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp \right) \right] 
\label{eq{5}}	    
\end{eqnarray}

@@

\subsection{Effect of multiple reflections on the delay spectrum} 
Correlation between the higher order reflections of the signal in any antenna with the direct component of the other would result in a delay response at integral multiple of $\Delta \tau$ and vice versa. 
If $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{eqnarray}
\volt_{1} & = &  a_{f} \volt_{sky}[1+ (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp + (1-\Gamma_{f})^{2} (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&&  ....+ (1-\Gamma_{f})^{n}(\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \nonumber
\end{eqnarray}
and,

\begin{eqnarray}
\volt_{2} & = &  a_{f} \volt_{sky}[1+ (1-\Gamma_{f})\Gamma_{f}\Gamma_{d} \dfngexp + (1-\Gamma_{f})^{2} (\Gamma_{f}\Gamma_{d})^2  (\dfngexp)^{2}+ \nonumber \\
&& ....+ (1-\Gamma_{f})^{n}(\Gamma_{f}\Gamma_{d})^{n} (\dfngexp)^{n}] \ast \fngexp \nonumber
\end{eqnarray}

@@

\begin{eqnarray}
\Sigma(\vec b, \nu) & = & a_{f}  [ 1+ 2 \Re [(1-\Gamma_{f}) \Gamma_{f}\Gamma_{d} \dfngexp+ (1-\Gamma_{f})^{2}(\Gamma_{f}\Gamma_{d})^2(\dfngexp)^{2}+ \nonumber\\
&&  ....+(1-\Gamma_{f})^{n} (\Gamma_{f}\Gamma_{d})^{n}(\dfngexp)^{n} ] ]\\
& = & { a_{f}\over \Gamma_{f} } \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i\phi}]^{n}
   \nonumber\\
   \label{eq6}
\end{eqnarray}

@@

\begin{eqnarray}\label{eqn:ratio1}
{I_{meas} \over I_{sky} } & = & \Sigma(\vec b, \nu) \nonumber\\
  & = & { a_{f}\over \Gamma_{f} } \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i\phi}]^{n}
   \nonumber\\
\end{eqnarray}

The right hand side of this equation is the instrument response kernel of a single HERA element as well as a representative of the instrument response in the measured visibility data between a pair of antennas. Our goal is to measure this kernel as a function of frequency and determine its effect on the delay spectrum of the visibility.

%%%%%%%%%%%%%%%%%
\subsection{Method}
The instrument response kernel is measured by measuring the return loss of the antenna element at the output port of the feed using a vector network analyser. The vector network analyser transmits a broadband noise of known amplitude and phase via a $75ft$ cable up to the feed. While a part of this signal  is returns to the VNA post reflection from the feed due to mismatch between the cable and the antenna impedance resulting in the primary return loss, rest of the signal is radiated out by the antenna and illuminates the dish and radiated into the space. However, the reflected signal from the dish vertex returns to
the feed. This incident signal is now
reflected back and forth in between the feed and the dish much like the sky
signal reflection discussed previously.

If $I_{tr}$ is the signal transmitted by the VNA , $\Gamma_{f}$ is the feed reflection coefficient, $I_{r}$
is the power incident back on the feed after reflection from the dish then the reflected power, $I_{refl}$,
back into the VNA would be:
\begin{equation}\label{eqn:series2}
I_{refl} =  I_{r}A_{f}[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

Once again, we consider one reflection from the feed and its subsequent reflection from the dish as one reflection in total. Equation \ref{eqn:series2} is similar to Equation \ref{eqn:series1}, with different incident powers.

$I_{r}$ is the power that is incident back on the feed, which is just the feed radiated power reflected off the dish:

@@

Hence the total returned power $P_{ret}$, to the VNA  would be:

@@

  \begin{eqnarray}\label{eqn:ratio2}
 {I_{ret} \over I_{tr} } & = & \Gamma_{f}
  +  \Gamma_{d}(1-\Gamma_f)^{2} [e^{i\phi}+ \Gamma_{f}\Gamma_{d} e^{i2\phi}+  ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{i n\phi}]\nonumber\\
  & = & \Gamma_{f} + { (1-\Gamma_f)a_{f}\over \Gamma_{f} } \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i\phi}]^{n}
   \nonumber\\
\end{eqnarray}

@@

The VNA measures the magnitude and phase of the quantity ${I_{ret}\over I_{tr}}$
as a function of frequency as shown in Figure \ref{fig:freq}. The magnitude and phase of the complex return loss is measured between 50 to 500MHz over 1024 frequency channels which gives a frequency resolution of 0.44MHz. The delay resolution of the measurements is ?t = 2.22ns. In our measurement
set-up, the first reflection occurs at the feed $\Gamma_{f}$, so
$({I_{ret} \over I_{tr} }  - \Gamma_{f}) $ gives an estimate of the delay
spectrum of the sky signal. In delay domain, the relative signal strength at
zero delay represents the factor $\Gamma_{f}$ while the signal strength at any
other delay represents any delayed signal that enters the feed after being
reflected from the feed surroundings.

%\section{Methodology}{\label{sec:methods}}

%Our reflectometry measurements are made using a prototype HERA dish at NRAO in Green Bank, WV. The dish is a $14\,m$ diameter
%parabolic reflector structurally supported with 3 telephone poles. The
%reflective material is made up of wire mesh that is attached to PVC
%pipes, forming the parabolic shape of the dish. With the current iteration of
%the HERA dish, the feed consists of a PAPER dipole encased in a cylindrical cage
%encompassing the backplane. The PAPER feed and the backplane (which is aimed at
%preventing feed-to-feed interaction between neighboring dishes) is raised and
%lowered by a three-pulley system. The focal height of the dish is $4.5m$
%($\sim{14.76}$ft).

%The heights measured in our experiment were made from
%the balun to the top of the concrete hub. However, the focal height should be
%measured from the backplane to point where the wire mesh would intersect the
%concrete hub at the vertex of the dish. To account for this height discrepency,
%we add 1.96 ft to all of our measured heights. These are the heights quoted in
%the plots.[XXX not done yet but will be done]
%Our measurements are made with a FieldFox in VNA mode. In this mode, a pulse
%is generated in the FieldFox and sent through a $75ft$ $50\Omega$ cable that
%connects to the feed with a 4:1 passive balun. The magnitude and phase of the complex return loss is measured between 50 to $500MHz$ over 1024 frequency channels which gives a frequency resolution of 0.44MHz. The delay resolution of the measurements is $\Delta{t}=2.22ns$.
%In addition, we note that the round trip of a reflection from the feed to the dish
%is $9m$, which corresponds to a delay of $30ns$.

@@

%need to add in third : Antenna feed focal length 4.5 m which corresponds to a
%signal propagation delay of 0.3 ns. Hence, subsequent reflections from the dish
%vertex are expected to be at a delay which is integral multiple of 0.3 nS.
\begin{figure}[ht!]
\centering
\includegraphics[totalheight=0.4\textheight]{plots/frequency_amp_phase_fullbw.png}
\caption{Amplitude and phase of the measured return loss. Colored dashed lines
mark three different frequency bands: $140-160MHz$ (typical power spectrum bandwidth), $100-200MHz$ (PAPER bandwidth), and
$50-250MHz$ (HERA bandwidth).}
\label{fig:freq}
\end{figure}
%++++++++++++++++++++++++++++++++++++++++++++++++++++++

\subsection{Estimation of the delay spectrum of the instrument}
Frequency domain data, as shown in Figure \ref{fig:window}, is Fourier-transformed to compute the response of the system in the delay domain. The signal transmitted by the VNA is reflected first from the feed. In delay domain,  this appears at the zero delay bin. We subtract this value from the measured data and divide the the data by a factor $(1-\Gamma_{a})$ to derive the delay spectrum for the system operating in receiving mode during sky observation.   %Because the measured data is band-limited between $50$ to $500MHz$, it is analogous to multiply the measured data by a square window function. 
%Then in delay domain, the response would be convolved with a $sinc$ function. This may result in excess power at high
%delays due to the sidelobes of the $sinc$ function. Hence, appropriate windowing of the measured data set is necessary before taking the Fourier transform.
%We have chosen a Hamming window for our analysis, and its effectiveness of this
%window function compared to others is illustrated in Figure
%\ref{fig:window}.
